Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
QA till Ch Results

\begin{enumerate}
  \item Choose the best suiting implementations of the SHA-3 candidates.
  \item Write Python bindings, with unit tests for verification, to be able to
    use the functions from Python.
  \item Modify the Tahoe-LAFS code to utilize the SHA-3 functions.
  \item Create a method of distributing new code fast and easy to the test grid.
  \item Set up the test environment and choose test vectors.
\end{enumerate}

@@

While the official \ac{NIST} \ac{SHA}-3 second round submissions contain both
a reference version and optimized versions of the candidates, it does not
appear to be updated since the submission deadline on September 15, 2009. They
also do not contain any benchmarking utility which could be used to decide on
the best implementation for a particular platform or architecture. \ac{SUPERCOP}
on the other hand, provides both, in combination with results stating which
compiler arguments that gives the best performance. On the downside, the
\ac{SUPERCOP} \ac{API} only specifies the {\verb crypto_hash() } function and
not an {\verb update() } function which Tahoe-\ac{LAFS} is dependent on using.

@@

The chosen implementations are shown in Table \ref{tbl:sha3:implementations}.
Their compiler flags used, which are extracted from the \ac{SUPERCOP} results,
can be seen in Table \ref{tbl:sha3:compilerflags}. Compiler flags are described
in the GCC manual\footnote{GCC Manual \url{http://gcc.gnu.org/onlinedocs/}}.
Compiler flags used for every implementation which might affect performance are:

\begin{itemize}
    \item -fomit-frame-pointer
    \item -mssse3
    \item -pthread
    \item -fPIC
\end{itemize}

@@

The relative speeds of the implementations chosen compared to
the best version in \ac{SUPERCOP}, and relative speed of the benchmark compared
to other candidates, can bee seen in Table \ref{tbl:sha3:speedrelative}.
Worth mentioning is that the fastest SHA-256 implementation in \ac{SUPERCOP} is
cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
Tahoe-\ac{LAFS}. For the most part we have found working versions of all
candidates in \ac{SUPERCOP}, however there are some exceptions.

@@

\paragraph{JH} seems to be broken in every implementation with regard to
Criterion 2. For both the optimized versions found in \ac{SUPERCOP} and the
reference implementations in the \ac{NIST} submission package,
{\verb Hash("foofoo") } will not yield the same result as two consecutive calls
to {\verb Update("foo") }.  However the implementation in sphlib does fulfill
this requirement. We have contacted the author, Thomas Pornin, which confirms
this problem with the official JH candidate and that the author has been made
aware of the problem but has not released any updated version yet
\cite{s_pornin}.

@@

\paragraph{Simd vect128 implementation} does not by default conform to Criterion
4, because of a bug in the implementation. This has been confirmed by the author
who also supplied us with a patch to correct the bug (Gaëtan Leurent, personal
communication, November 10, 2010).

@@

Since Tahoe-\ac{LAFS} is written in Python, we needed some way of interacting
with the C implementations of the NIST candidates. We did this by the use of
Cython, and built a Python object-oriented library which we named SHA3lib.
SHA3lib supports all of the 14 second round \ac{SHA}-3 candidates, but only in
their 256-bit form.

\subsection{SHA3lib}
SHA3lib is built to mimic the functionality of
hashlib\footnote{\url{http://docs.python.org/library/hashlib.html}}, the
default Python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. This
made us create class names for each candidate on the form
\emph{candidate name+256}. In addition, the classes included the following
functions:

\begin{itemize}
    \item Constructor with an optional data element that does the same as
    {\verb update() }
    \item {\verb update() } takes data that should be considered in hashing
    \item {\verb digest() } that produce a raw byte-digest from submitted data
    \item {\verb hexdigest() } that produce a hex representation of the digest
    \item {\verb copy() } that copies the internal state of the function and
        produces a copy of the object.
\end{itemize}

\subsubsection{Design}
Our original goal was to make SHA3lib include functionality to support all
output lengths, and not only 256-bit. This was based on our observation that
the official \ac{NIST} submission packages include ``all-in-one''
implementations that support this. Later we observed that this is not the case
for all implementations in \ac{SUPERCOP}. Together with the fact that
Tahoe-\ac{LAFS} uses 256-bit output we in the end only included support for
256-bit output in SHA3lib. The design we have used is shown in Figure
\ref{fig:python:inheritance}, and does follow the initial idea that SHA3lib
should support multiple output lengths.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.6\columnwidth]{python-inheritance.pdf}
    \caption{Design of SHA3lib using echo as an example}
    \label{fig:python:inheritance}
\end{figure}

Basically the class \textbf{echo} resides in Cython and in C, and is responsible
for bringing the C functionality of the C implementation to Python. By compiling
\textbf{echo} it will be made available for use in Python.  The class
\textbf{echo256} is a simple Python class which inherits from \textbf{echo}. The
only action this class actually takes is to set the output size to 256-bit. If
the underlying C-functions would have supported multiple output sizes the same
could have been done to incorporate those.

\subsubsection{Cython}
We employ Cython to generate bindings for the C implementation and the
different candidates, as well as to create the base Python classes. To access
functions in C we need to generate a Cython header file, which pretty much does
the same as a C header file; it tells which functions and data types are
available. How we accomplish this for an implementation that satisfies the
\ac{NIST} \ac{API} is shown in Listing \ref{lst:cython_header}.

@@

We then create a wrapper-class that mimics hashlib, this is shown in Listing
\ref{lst:cython_impl}. Functions declared with def er Python functions,
{\verb cdef } are Cython only while {\verb cpdef } are functions that can be
accessed from both. While this is just the implementation for the echo
candidate, the almost exact same pattern is used for all of the other
candidates.

@@

\subsubsection{Distutils}
We use \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
build and install our code. The alternative to doing this would be to manually
transform Cython files to C-files, build the shared library from this and other
relevant C-files and then install by copying into the correct system directory
where Python libraries should be stored. \ac{Distutils} also support expansion
through self defined commands, something we have utilised to do Unit Testing.
Options and configuration for \ac{Distutils} is done through a file
{\verb setup.py}. Relevant configuration for compiling the echo extension, and
installing SHA3lib can bee seen in Listing \ref{lst:setup_py}.

@@

\paragraph{Known-Answer Tests} are used to check that the implementation conform
to the specifications of the algorithm author. The \ac{KAT}s include
bytestrings, their length and their expected message digest. The length of the
inputs range from 0 bits to 34304 bits, and also include a message that should
be repeated for about 17 million times, to verify that the implementation can handle
large inputs.

\paragraph{Basic Hash Tests} is what we have named the set of tests which test
the core basics of what a hash function implementation will have to be able to
do to be usable. This include that Init(), Update() and Final() works as
expected, and that the output is of correct size. This include that every
function should produce the same digest for the inputs Update("foofoo") and two
consecutive calls to Update("foo").

\paragraph{Copy Hash Tests} does not test any of the functionality used by
Tahoe-\ac{LAFS}, but the copy()-function is a part of Hashlib, and we have
included it in SHA3lib to conform to this. The idea is that if we have two
bytestring that share a common prefix, we can input this prefix to Update(),
then copy the state. The tests check if this is possible, and that altering the
copied state does not affect the original.

@@

The code of Tahoe-LAFS conforms to the style guide of Python, known as PEP 8.
This, along with the code being highly modular and consistent to its
architecture, emphasizes code readability.
With some experience using Python, it is not difficult to learn the Tahoe code
enough to understand what is happening when a file is uploaded or downloaded
with the system.

@@

As stated in Section \ref{sec:lengthextension}, none of the SHA-3 candidates is
vulnerable to the length-extension attack. Since we are interested in finding
out how Tahoe-\ac{LAFS} performs when utilizing the SHA-3 functions, we can
remove the lines that fixes the length-extension problems manually. In addition,
to try out one of the new implementations, we change line one in Listing
\ref{lst:hashutil}, and an example of this can be seen in Listing
\ref{lst:hashutilmod}.

@@

For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
changing this line, because they contain test fixtures made with the included
SHA-256.  An implication of this is that uploading of mutable files renders
impossible, since the internal verification fails. For this reason, \emph{only
immutable files} are tested and measured in this paper. Implications of this are
discussed further in Chapter \ref{ch:discussion}.

\section{Automatic testing and distribution}
We fully automate the testing procedure. This is achieved by the use of
software revision control systems, which distributes both the source code of
Tahoe-LAFS, SHA3lib and shell-scripts for installation and configuration of
both. Configuration is mainly done by the use of regular expressions.

@@

\vspace{12pt}
\noindent For every \ac{SHA}-3 candidate:
        \begin{enumerate}
            \item Install SHA3lib on all nodes
            \item Replace used hash algorithm in Tahoe-LAFS with current
            candidate
            \item Install Tahoe-LAFS
            \item Create and start one Tahoe-LAFS introducer
            \item Create and start Tahoe-LAFS storage nodes on 3 nodes
            \item Create a Tahoe-LAFS client on a previously unnused node
            \item Do testing
            \item Clean up
        \end{enumerate}

\paragraph{Distribution of source code for Tahoe-LAFS and SHA3lib} is done by
\ac{SVN} and GIT respectivly. The change of hash algorithm in Tahoe-LAFS is
however done with the use of a regular expression, because of both the
simplicity of the change, and to not clutter the revision control system. Both
Tahoe-LAFS and SHA3lib uses \ac{Distutils}, which includes options to install
them globally on a system.

@@

During these operations, we counted the number of hashing operations performed,
and timed total time spent hashing, i.e. the total time it took for the function
calls to the underlying hashing operations to complete.

>>>>

\begin{enumerate}
  \item Choose the best suiting implementations of the SHA-3 candidates.
  \item Write Python bindings, with unit tests for verification, to be able to
    use the cryptographic hash functions from Python.
  \item Modify the Tahoe-LAFS code to utilize the SHA-3 functions.
  \item Create a method of distributing new code fast and easy to the test grid.
  \item Set up the test environment and choose test vectors.
\end{enumerate}

@@

While the official \ac{NIST} \ac{SHA}-3 second round submissions contain both
a reference version and optimized versions, they do not
appear to be updated since the submission deadline on September 15, 2009. They
do also not contain any benchmarking utility which could be used to decide on
the best implementation for a particular platform or architecture. \ac{SUPERCOP}
on the other hand, provides both, in combination with results stating which
compiler arguments giving the best performance. On the downside, the
\ac{SUPERCOP} \ac{API} only specifies the {\verb crypto_hash() } function and
not an {\verb update() } function which Tahoe-\ac{LAFS} is dependent on using.

@@

The chosen implementations are shown in Table \ref{tbl:sha3:implementations}.
Their compiler flags used, which are extracted from the \ac{SUPERCOP} results,
can be seen in Table \ref{tbl:sha3:compilerflags}. Compiler flags are described
in the GCC manual\footnote{GCC Manual \url{http://gcc.gnu.org/onlinedocs/}},
and the one used for every implementation which might affect performance, are:

\begin{itemize}
    \item[] -fomit-frame-pointer
    \item[] -mssse3
    \item[] -pthread
    \item[] -fPIC
\end{itemize}

@@

The relative speeds of the implementations chosen compared to
the best version in \ac{SUPERCOP}, and relative speed of the benchmark compared
to other candidates, can bee seen in Table \ref{tbl:sha3:speedrelative}.
Worth mentioning is that the fastest SHA-256 implementation in \ac{SUPERCOP} is
from cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
Tahoe-\ac{LAFS}. For the most part, we have found working versions of all
candidates in \ac{SUPERCOP}, however there are some exceptions.

@@

\paragraph{JH} seems to be broken in every implementation with regard to
Criterion 2. For both the optimized versions found in \ac{SUPERCOP}, and the
reference implementations in the \ac{NIST} submission package,
{\verb Hash("foofoo") } will not yield the same result as two consecutive calls
to {\verb Update("foo") }. However, the implementation in sphlib does fulfill
this requirement. We have contacted the author, Thomas Pornin, which confirms
this problem with the official JH candidate, and that the author has been made
aware of the problem, but has not released any updated version yet
\cite{s_pornin}.

@@

\paragraph{Simd vect128} implementation does not by default conform to Criterion
4, because of a bug in the implementation. This has been confirmed by the author
who also supplied us with a patch to correct the bug \cite{s_simd}.

@@

Since Tahoe-\ac{LAFS} is written in Python, we needed some way of interacting
with the C implementations of the \ac{NIST} candidates. We did this by the use of
Cython, and built a Python object-oriented library which we named SHA3lib.
SHA3lib supports all of the 14 second round \ac{SHA}-3 candidates, but only in
their 256-bit form.

\subsection{SHA3lib}

SHA3lib is built to mimic the functionality of
hashlib\footnote{\url{http://docs.python.org/library/hashlib.html}}, the
default Python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. This
made us create class names for each candidate on the form
\emph{candidate name+256}. In addition, the classes included the following
functions:

\begin{itemize}
    \item Constructor with an optional data element that does the same as
    {\verb update() }
    \item {\verb update() } receives data that should be hashed
    \item {\verb digest() } that produce a raw byte digest from submitted data
    \item {\verb hexdigest() } that produce a hex representation of the digest
    \item {\verb copy() } that copies the internal state of the function and
        produces a copy of the object.
\end{itemize}

\subsubsection{Design}

Our original goal was to make SHA3lib include functionality to support all
output lengths, and not only 256-bit. This was based on our observation that the
official \ac{NIST} submission packages include ``all-in-one'' implementations
that support this. Later we observed that this is not the case for all
implementations in \ac{SUPERCOP}. Together with the fact that Tahoe-\ac{LAFS}
uses 256-bit output, in the end only included support for 256-bit output in
SHA3lib was implemented. The design we have used is shown in Figure
\ref{fig:python:inheritance}, and does follow the initial idea that SHA3lib
should support multiple output lengths.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.6\columnwidth]{python-inheritance.pdf}
    \caption{Design of SHA3lib using Echo as an example}
    \label{fig:python:inheritance}
\end{figure}

Basically the class \textbf{echo} resides in Cython and in C, and is responsible
for bringing the C functionality of the C implementation to Python. By compiling
\textbf{echo} it will be made available for use in Python.  The class
\textbf{echo256} is a simple Python class which inherits from \textbf{echo}. The
only action this class actually performs, is to set the output size to 256 bit. If
the underlying C-functions would have supported multiple output sizes, the same
could have been done to incorporate those.

\subsubsection{Cython}
We employ Cython to generate bindings for the C implementation of the
different candidates, as well as to create the base Python classes. To access
functions in C we need to generate a Cython header file, which pretty much does
the same as a C header file -- it describes which functions and data types are
available. How we accomplish this for an implementation that satisfies the
\ac{NIST} \ac{API}, is shown in Listing \ref{lst:cython_header}.

@@

We then create a wrapper-class that mimics hashlib, this is shown in Listing
\ref{lst:cython_impl}. Functions declared with {\verb def } are Python
functions, {\verb cdef } are Cython-only functions, while {\verb cpdef } are
functions that can be accessed from both languages. While this is just the
implementation for the Echo candidate, the almost exact same pattern is used for
all of the other candidates.

@@

\subsubsection{Distutils}

We use \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
build and install our code. The alternative of doing this would be to manually
transform Cython files to C files, build the shared library from this and other
relevant C files, and then install by copying into the correct system directory
where Python libraries should be stored. \ac{Distutils} also support expansion
through self defined commands, something we have utilised to do Unit Testing.
Options and configuration for \ac{Distutils} is done through a file named
{\verb setup.py }. Relevant configuration for compiling the Echo extension, and
installing SHA3lib, can bee seen in Listing \ref{lst:setup_py}.

@@

\paragraph{Known-Answer Tests} are used to check that the implementation conform
to the specifications of the algorithm author. The \acp{KAT} include
byte strings, their length and their expected message digest. The length of the
inputs range from 0 bits to 34304 bits, and also include a message that should
be repeated for about 17 million times, to verify that the implementation can handle
large inputs.

\paragraph{Basic Hash Tests} are what we have named the set of tests which test
the core basics of what a hash function implementation will have to be able to
do to be usable. This include that {\verb init() }, {\verb update() } and
{\verb final() } works as expected, and that the output is of correct size. This
include that every function should produce the same digest for the inputs
{\verb update("foofoo") } and two consecutive calls to {\verb update("foo") }.

\paragraph{Copy Hash Tests} do not test any of the functionality used by
Tahoe-\ac{LAFS}, but the {\verb copy() } function is a part of hashlib, and we
have included it in SHA3lib to conform to this. The idea is that if we have two
byte strings that share a common prefix, we can input this prefix to
{\verb update() }, then copy the state. The tests check if this is possible, and
that altering the copied state does not affect the original.

@@

The code of Tahoe-\ac{LAFS} conforms to the style guide of Python, known as PEP
8.  This, along with the code being highly modular and consistent to its
architecture, emphasizes code readability.  With some experience using Python,
it is not difficult to learn the Tahoe code enough to understand what is
happening when a file is uploaded or downloaded with the system.

@@

As stated in Section \ref{sec:lengthextension}, none of the \ac{SHA}-3
candidates are vulnerable to the length-extension attack. Since we are
interested in finding out how Tahoe-\ac{LAFS} performs when utilizing the
\ac{SHA}-3 functions, we can remove the lines that fixes the length-extension
problems manually. In addition, to try out one of the new implementations, we
change line one in Listing \ref{lst:hashutil}, and an example of this can be
seen in Listing \ref{lst:hashutilmod}.

@@

For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
changing this line, because they contain test fixtures made with the included
SHA-256.  An implication of this is that uploading of mutable files renders
impossible, since the internal verification fails. For this reason, \emph{only
immutable files} are tested and measured in this paper. Implications of this are
further discussed in Chapter \ref{ch:discussion}.

\section{Automatic testing and distribution}
We fully automate the testing procedure. This is achieved by the use of
software revision control systems, which distributes both the source code of
Tahoe-\ac{LAFS}, SHA3lib and shell-scripts for installation and configuration of
both. Configuration is mainly done by the use of regular expressions.

@@

\vspace{12pt}
\noindent For every \ac{SHA}-3 candidate:
        \begin{enumerate}
            \item Install SHA3lib on all nodes
            \item Replace used hash algorithm in Tahoe-\ac{LAFS} with current
              candidate
            \item Install Tahoe-\ac{LAFS}
            \item Create and start one Tahoe-\ac{LAFS} introducer
            \item Create and start Tahoe-\ac{LAFS} storage nodes on 3 nodes
            \item Create a Tahoe-\ac{LAFS} client on a previously unused node
            \item Do testing
            \item Clean up
        \end{enumerate}

\paragraph{Distribution of source code for Tahoe-LAFS and SHA3lib} are done by
\ac{SVN} and GIT respectively. The change of hash algorithm in Tahoe-LAFS is
however done with the use of a regular expression, because of both the
simplicity of the change, and to not clutter the revision control system. Both
Tahoe-LAFS and SHA3lib uses \ac{Distutils}, which includes options to install
them globally on a system.

@@

During these operations, we counted the number of hashing operations performed,
and timed total time spent hashing, i.e. the total time it took for the function
calls to the underlying hashing operations to complete. We also timed how long
the Tahoe operation took in total for one test vector.
